```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
#install.packages("ggplot2")
library(ggplot2)
require(ggplot2)
#install.packages('gridExtra')
library(gridExtra)
```



```{r, include=FALSE}
setwd("/Users/Remi/Dropbox/1_WORK/2015/4_PROJECTS/1_SOFTWARE/IMAGEJ_LIB/CELL SIGNALING ANALYSIS/CODE/6_DATA_CLEANING")
```
# Data Cleaning: TECHNICAL and CONSISTENCY corrections of the data
Below is listed the standard workflow I run for data cleaning of the extracted data from the image acquisitions  


i) Show all attributes names

ii) Check that the number of columns of numerical attributes is consistent with what is expected

iii) Show whether there is any column with special characters (NA, NaN, Null, Inf, -Inf) 

iv) Change found special caracters with a specific numeircal value

v) Check whether there are inconsistent attributes values, e.g,  negative values



```{r}
#Load the raw data
DT <- read.delim("RAW_DATA_SAMPLE.txt")

#print(paste('total number of attributes: ', dim(DT)[2]))
#print(paste('total number of rows: ', dim(DT)[1]))
```

```{r echo=FALSE}
print(paste('total number of attributes: ', dim(DT)[2]))
print(paste('total number of rows: ', dim(DT)[1]))
```


### i) Show all attributes names
```{r}

# SHOWS THE ATTRIBUTES NAMES (the first 5 attributes should  contain textual information)
print(names(DT))

```

### ii) Check that the number of columns of the numerical attributes is consistent with what is expected
```{r}

# INPUT DATA
idx_first_numeric_attribute = 4 # (skip the first attributes related to textual information)
num_image_channels = 4
num_numerical_attributes_channels=33

# total number of numerical attributes
tot_num_numerical_cols = num_image_channels * num_numerical_attributes_channels

# count the numerical and non-numerical attributes
dm = dim(DT)
num_cols = dm[2]
count_numeric_attributes = 0;
no_correct_attributes = c('')
no_correct_idx_attributes = c()
count_no_correct = 0;
# report the columns with elements to correct
for(i in idx_first_numeric_attribute:num_cols){
  
  if(class(DT[,i])=='numeric' || class(DT[,i])=='integer' )
    count_numeric_attributes = count_numeric_attributes + 1 
  else{
    count_no_correct = count_no_correct + 1
    no_correct_attributes[count_no_correct] = names(DT[i])
    no_correct_idx_attributes[count_no_correct] = i
  }
}

# outcome
if(count_no_correct==0 && (tot_num_numerical_cols == count_numeric_attributes)){
  print('THE NUMBER OF ATTRIBUTES IS CORRECT')
  print(paste('total number of found numerical attribute: ',count_numeric_attributes ))
  print(paste('total number of expected numerical attribute: ',tot_num_numerical_cols ))
} else{
  print('THE NUMBER OF ATTRIBUTES IS NOT CORRECT')
  print(paste('total number of found numerical attribute: ',count_numeric_attributes  ))
  print(paste('total number of expected numerical attribute: ', tot_num_numerical_cols))
  print('non-numerical attributes names: ' )
  print(no_correct_attributes)
  print('non-numerical attributes indexes:')
  print(no_correct_idx_attributes)
}
```

### iii) Show whether there are columns with special characters (NA, NaN, Null, Inf, -Inf) 


```{r}
# total number of attributes/columns
num_cols = dim(DT)[2]

# report the columns with elements to correct
count_numeric_attributes=0
found_attributes = c('')

for(i in 1:num_cols){
    a = is.finite(DT[,i])
    if(length(a[a==FALSE])>0){
         count_numeric_attributes = count_numeric_attributes +1
        found_attributes[count_numeric_attributes] = names(DT[i])
    }
}

print(paste('Number of found attributes with special characters:', count_numeric_attributes))
print(found_attributes)
```

### iv) Change the found special characters values with a given numerical value
```{r}
# total number of attributes/columns
num_cols = dim(DT)[2]

# report the columns with elements to correct
for(i in 1:num_cols){
   if(class(DT[,i])=='numeric' || class(DT[,i])=='integer'){
      a = is.finite(DT[,i])
      DT[,i][a==FALSE]=0
   }
}
```

### v) Check there are attribute with inconsistent values, e.g,  negative values
```{r}
# total number of attributes/columns
num_cols = dim(DT)[2]

# report the columns with elements to correct
count_numeric_attributes=0
found_attributes = c('')

for(i in 1:num_cols){
  
    if(class(DT[,i])=='numeric' || class(DT[,i])=='integer' ){
      a = (DT[,i]<0 && is.finite(DT[,i]))
      if(length(a[a==TRUE])>0){
           count_numeric_attributes = count_numeric_attributes +1
          found_attributes[count_numeric_attributes] = names(DT[i])
      }
    }
}

print(paste('Number of found attributes with negative values:', count_numeric_attributes))
print(found_attributes)

```

# SAve the cleaned data frame
```{r}
write.table(DT,file="RAW_DATA_SAMPLE_CLEANED.txt", sep="\t")
```

